### 2014-12-02 writing/F'14 paper/abstract.tex
Paper update.

%%%% POTTS REVISION
\begin{abstract} 
  Recursive neural networks (RNNs) for sentence meaning have been
  successful in many tasks, but it remains an open question whether
  any such model can learn compositional semantic representations that
  support logical deduction. We pursue this question by
  evaluating whether two such models---RNNs and recursive neural
  tensor networks (RNTNs)---can correctly learn to identify logical
  relationships such as entailment and contradiction. % between sentences.
  In our first set of experiments, we generate artificial data from a
  logical grammar and use it to evaluate the models' ability to learn
  to handle basic relational reasoning, recursive structures, and
  quantification. We then evaluate the models on the more naturalistic
  SICK challenge data. Both models perform competitively on the SICK
  data and generalize well in all three experiments on simulated data,
  suggesting that they can learn suitable representations for logical
  inference in natural language. 
\end{abstract}

% TODO: Is there anything I should add that would be more specific than 'many tasks'? 'Tasks like paraphrase detection'?

>>>>

%%%% POTTS REVISION
\begin{abstract} 
  Recursive neural networks (RNNs) for sentence meaning have been
  successful for many tasks, but it remains an open question whether
  any such model can learn compositional semantic representations that
  support logical deduction. We pursue this question by
  evaluating whether two such models---RNNs and recursive neural
  tensor networks (RNTNs)---can correctly learn to identify logical
  relationships such as entailment and contradiction. % between sentences.
  In our first set of experiments, we generate artificial data from a
  logical grammar and use it to evaluate the models' ability to learn
  to handle basic relational reasoning, recursive structures, and
  quantification. We then evaluate the models on the more natural
  SICK challenge data. Both models perform competitively on the SICK
  data and generalize well in all three experiments on simulated data,
  suggesting that they can learn suitable representations for logical
  inference in natural language. 
\end{abstract}

### 2014-12-02 writing/F'14 paper/discussion.tex
Paper update.

\section{General discussion}\label{sec:discussion}

This paper evaluated two recursive models on a series of three increasingly
challenging interpretive tasks involving natural language inference---
the core relational algebra of natural logic with entailment and
exclusion; recursive propositional logic structures; and statements
involving quantification and negation---as well as on an entailment task on
natural language image captions. The results suggest that RNTNs,
but not necessarily plain RNNs, have the capacity to meet the 
challenges of these tasks with reasonably-sized training sets. These positive results are
promising for the future of learned representation models in the
applied modeling of compositional semantics.

Of course, challenges remain. In terms of our experimental data, even
the RNTN falls short of perfection in the recursion experiment, with
performance falling off steadily as the size of the expressions grows. It
remains to be seen whether these deficiencies can be overcome with
stronger models or the optimization procedures. The space of possible stronger
models is sizable, including extensions of the recursive models used here
\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, recurrent
ones \cite{sutskever2014sequence}, and even learning-oriented variants 
of more structured models like those of \newcite{grefenstette2013towards} or \newcite{rocktaschellow}.
In addition,
there remain subtle questions about how to fairly assess whether these
models have truly generalized in the way we want them to. There is a
constant tension between showing the models training data that gives
them a chance to learn the target logical functions and revealing the
answer to them in a way that leads to overfitting. The underlying
logical theories provide only limited guidance on this point.

Finally, our artificial data experiments have only scratched the surface 
of the logical complexity of natural language; in future experiments, we hope to test sentences
with embedded quantifiers, multiple interacting quantifiers, relative
clauses, and other kinds of recursive structure. Nonetheless, the
rapid progress the field has made with these models in recent years
provides ample reason to be optimistic that they can be trained to
meet the challenges of natural language semantics.

>>>>

\section{Discussion and conclusion}\label{sec:discussion}

This paper first evaluates two recursive models on a series of three increasingly
challenging interpretive tasks involving natural language inference---the 
core relational algebra of natural logic with entailment and
exclusion; recursive propositional logic structures; and statements
involving quantification and negation. We then showed that the same models can learn to
perform an entailment task on natural language image captions. \mynote{The results suggest that RNTNs,
but not necessarily plain RNNs}, have the capacity to model these tasks with 
reasonably-sized training sets. These positive results are
promising for the future of learned representation models in the
applied modeling of compositional semantics.

Of course, challenges remain. In terms of our experimental data, even
the RNTN falls short of perfection in the recursion experiment, with
performance falling off steadily as the size of the expressions grows. It
remains to be seen whether these deficiencies can be overcome with
stronger models or optimization procedures. The space of possible stronger
models is sizable, including extensions of the recursive models used here
\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, recurrent
models \cite{sutskever2014sequence}, and even learning-enabled variants 
of more structured models like those of \newcite{grefenstette2013towards} or \newcite{rocktaschellow}.
In addition,
there remain subtle questions about how to fairly assess whether these
models have truly generalized in the way we want them to. There is a
constant tension between showing the models training data that gives
them a chance to learn the target logical functions and revealing the
answer to them in a way that leads to overfitting. The underlying
logical theories provide only limited guidance on this point.

Our artificial data experiments have only scratched the surface 
of the logical complexity of natural language; in future experiments, we hope to test sentences
with embedded quantifiers, multiple interacting quantifiers, relative
clauses, and other kinds of recursive structure. Nonetheless, the
rapid progress the field has made with these models in recent years
provides ample reason to be optimistic that they can be trained to
meet all the challenges of natural language semantics.

### 2014-12-02 writing/F'14 paper/intro.tex
Paper update.

To date, experimental investigations of these concepts using
distributed representations have been largely confined to short
phrases
\cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment}.
However, for robust natural language understanding, it is essential to
model these phenomena in their full generality on complex linguistic
structures. Other recent work
\cite{grefenstette2013towards,rocktaschellow} has developed rich
frameworks for computing the meanings of complex fragments of natural
language compositionally from vector or tensor representations, but
these approaches do not confront the question of how to learn their
highly structured representations from data using existing applied
machine learning models.

% TODO: Any way to avoid nested parens? Yes, use \citealt ---CP

@@

The experiments with simulated data provide a convincing demonstration
of the ability of neural networks to learn to encode semantic
inferences in complex natural language sentences from reasonably-sized
training sets. However, we are also interested in the more practical
question of whether they can learn these representations from more
naturalistic text. To address this question, we apply our models to
the SICK entailment challenge data. The small size of this corpus puts
data-intensive neural models like ours at a disadvantage, but we are
nonetheless able to achieve competitive performance on it
\mynote{CONSIDER ADDING A CONCRETE DETAIL --- E.G., ACCURACY
  AND ITS RELATION TO THE COMPETITION MEDIAN}, suggesting that the
representational abilities that we observe in the previous sections
are not limited to carefully circumscribed tasks.

>>>>

To date, experimental investigations of these concepts using
distributed representations have been largely confined to short
phrases
\cite{Mitchell:Lapata:2010,Grefenstette-etal:2011,baroni2012entailment}.
However, for robust natural language understanding, it is essential to
model these phenomena in their full generality on complex linguistic
structures. Other recent work
\cite{grefenstette2013towards,rocktaschellow} has developed rich
frameworks for computing the meanings of complex fragments of natural
language compositionally from vector or tensor representations, but
these approaches have not yet attempted to learn their
highly structured representations from data using existing applied
machine learning models.

@@

The experiments with simulated data provide a convincing demonstration
of the ability of neural networks to learn to encode semantic
inferences in complex natural language sentences from reasonably-sized
training sets. However, we are also interested in the more practical
question of whether they can learn these representations from more
naturalistic text. To address this question, we apply our models to
the SICK entailment challenge data. The small size of this corpus puts
data-intensive neural models like ours at a disadvantage, but we are
nonetheless able to achieve competitive performance on it
\mynote{CONSIDER ADDING A CONCRETE DETAIL --- E.G., ACCURACY
  AND ITS RELATION TO THE COMPETITION MEDIAN}, suggesting that the
representational abilities that we observe in the previous sections
are not limited to carefully circumscribed tasks. We conclude that RNTN models
are adequate for typical cases of natural logic inference, and that there is not yet
any clear level of inferential complexity for which other approaches work and NN models fail.

% TODO: Clean up and condense to column length?

### 2014-12-02 writing/F'14 paper/join.tex
Minor paper updates.
Paper update.

\paragraph{Experiments}
To test our models' ability to learn these inferential patterns, we
create small boolean structures for our logic in which terms denote
sets of entities from a small domain.  Figure~\ref{lattice-figure}
depicts a structure of this form. The lattice gives the full model,
for which all the statements on the right are valid. We divide these
statements evenly into training and test sets, and remove from the
test set those statements which cannot be proven from the training
statements.
% using the logic depicted in Figure~\ref{lattice-figure}.

@@

We trained versions of both the RNN model and the RNTN model on these
data sets. In both cases, the models were implemented exactly as
described in Section~\ref{methods}, but since the items being compared
are single terms rather than full tree structures, the composition
layer was not involved, and the two models differed only in which
layer function was used for the comparison layer. We simply present
the models with two single terms, each of which corresponds to a
single vector in the (randomly initialized) vocabulary matrix $V$,
thereby ensuring that the model has no information about the terms
being compared except the relations between them.

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{??d RNN}  & \multicolumn{2}{c}{??d RNTN}\\
    \midrule
    Provable test &52.0 &(10.3) &	95...&(-)& \textbf{96.8...} & \textbf{(-)}\\
    All test &53.8 &(10.5) &	-&(-)& \textbf{-} & \textbf{(-)}\\
    Train &53.8 &(10.5)    & -&(-)& \textbf{-} & \textbf{(-)}\\

@@

% TODO: Do we still care about the unprovable test examples?

\paragraph{Results} 
Table \ref{joinresultstable} shows the results, which are fairly straightforward 
to interpret. Both models were able to accurately encode the relations 
between the terms in the geometric relations between their vectors, 
and was able to then use that information to recover relations that 
were not overtly included in the training data.  \tbchecked{The model's slightly 
poorer performance on the unprovable test examples shows that it
 has not simply encoded an approximation the entire underlying 
 model, but is rather taking advantage of the specific statements seen in the test data.}

>>>>

\paragraph{Experiments}
To test our models' ability to learn these inferential patterns, we
create small boolean structures for our logic in which terms denote
sets of entities from a small domain.  Figure~\ref{lattice-figure}
depicts a structure of this form. The lattice gives the full model,
for which all the statements on the right are valid. We divide these
statements evenly into training and test sets, and remove from the
test set statements which cannot be proven from the training
statements.
% using the logic depicted in Figure~\ref{lattice-figure}.

@@

We trained versions of both the RNN model and the RNTN model on these
data sets. In both cases, the models are implemented exactly as
described in Section~\ref{methods}, but since the items being compared
are single terms rather than full tree structures, the composition
layer is not involved, and the two models differed only in which
layer function is used for the comparison layer. We simply present
the models with two single terms, each of which corresponds to a
single vector in the (randomly initialized) vocabulary matrix $V$,
thereby ensuring that the model has no information about the terms
being compared except the relations between them.

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{??d RNN}  & \multicolumn{2}{c}{??d RNTN}\\
    \midrule
    Test &52.0 &(10.3) &	95...&(-)& \textbf{96.8...} & \textbf{(-)}\\
    Train &53.8 &(10.5)    & -&(-)& \textbf{-} & \textbf{(-)}\\

@@

\paragraph{Results} 
Table \ref{joinresultstable} shows the results, which are straightforward 
to interpret. Both models are able to accurately encode the relations 
between the terms in the geometric relations between their vectors, 
and is able to then use that information to recover relations that 
are not overtly included in the training data.

### 2014-12-02 writing/F'14 paper/methods.tex
Paper update.

We study neural models that adhere to the linguistic \ii{principle of
 compositionality}, which says that the meanings for complex
expressions are derived from the meanings of their constituent parts
via specific composition functions \cite{Partee84,Janssen97}. In our
distributed setting, word meanings are vectors of dimension $n$. The
composition function maps pairs of them to single vectors of dimension $n$, 
which can then be merged again to represent more complex
phrases. Once the entire sentence-level representation has been
derived, it serves as a fixed-dimensional input for some subsequent function.

We use the model architecture depicted in Figure~\ref{sample-figure}. 
The two phrases being compared are processed separately using a pair 
of recursive networks that share a single set of parameters. 
The resulting vectors are fed into a separate comparison
layer (fixed at 75 dimensions) that is meant to generate a feature vector capturing the
relation between the two phrases. The output of this layer is then
given to a softmax classifier, which in turn produces a hypothesized
distribution over the seven relations represented in Table~\ref{b-table}.

@@

The comparison layer uses the same layer function as the
composition layers (either an NN layer or an NTN layer) with
independently learned parameters and a separate nonlinearity function.
Rather than use a $\tanh$ nonlinearity here, we found better results with a rectified linear function. In
particular, we use the leaky rectified linear function
\shortcite{maasrectifier}: $f(x)=\max(x, 0) +
0.01\min(x, 0)$.

@@

% Removed 'minibatch' -> For two of the experiments, we use minibatches
% of size 1, which doesn't really count.
We train the model using stochastic gradient descent (SGD)
with learning rates computed using AdaDelta \cite{zeiler2012adadelta}.
The parameters (including the vocabulary) are initialized randomly.
Because the classes are not balanced in general, we report performance
using both accuracy and macroaveraged F1, since the latter emphasizes
 performance on infrequent classes. We compute macroaveraged F1 
as the harmonic mean of mean precision and mean recall, both computed
for all classes for which there is test data, setting precision to 0 
where it is not defined. Source code and generated data will be released
after the review period.

>>>>

We study neural models that adhere to the linguistic \ii{principle of
 compositionality}, which says that the meanings for complex
expressions are derived from the meanings of their parts
via specific composition functions \cite{Partee84,Janssen97}. In our
distributed setting, word meanings are vectors of dimension $n$. The
composition function maps pairs of them to single vectors of dimension $n$, 
which can then be merged again to represent more complex
phrases. Once the entire sentence-level representation has been
derived, it serves as a fixed-dimensional input for some subsequent function.

We use the tree pair model architecture depicted in Figure~\ref{sample-figure}, which is novel 
but closely modeled after past work on recursive NNs. 
The two phrases being compared are processed separately using a pair 
of recursive networks that share a single set of parameters. 
The resulting vectors are fed into a separate comparison
layer (fixed at 75 dimensions) that is meant to generate a feature vector capturing the
relation between the two phrases. The output of this layer is then
given to a softmax classifier, which in turn produces a hypothesized
distribution over the seven relations represented in Table~\ref{b-table}.

@@

The comparison layer uses the same layer function as the
composition layers (either an NN layer or an NTN layer) with
independently learned parameters and a separate nonlinearity function.
Rather than use a $\tanh$ nonlinearity here, we found better results with a rectified linear function. In
particular, we use the leaky rectified linear function
\cite{maasrectifier}: $f(x)=\max(x, 0) +
0.01\min(x, 0)$.

@@

% Removed 'minibatch' -> For two of the experiments, we use minibatches
% of size 1, which doesn't really count.
We train the model using stochastic gradient descent (SGD)
with learning rates computed using AdaDelta \cite{zeiler2012adadelta}.
The parameters (including the vocabulary) are initialized randomly.
Because the classes are not balanced in general, we report performance
using both accuracy and macroaveraged F1, since the latter emphasizes
 performance on infrequent classes. We compute macroaveraged F1 
as the harmonic mean of average precision and average recall, both computed
for all classes for which there is test data, setting precision to 0 
where it is not defined. Source code and generated data will be released
after the review period.

### 2014-12-02 writing/F'14 paper/quantifiers.tex
Paper TODO tweaks
Minor paper updates.
Paper update.

% TODO: Compress?

We have seen that recursive NNs can learn an approximation of propositional
logic.  However, natural languages can express functional meanings of
considerably greater complexity than this.  As a first step towards
investigating whether our models can capture this complexity, we now
attempt to directly measure the degree to which they are able to
develop suitable representations for the semantics of natural language
quantifiers like \ii{some} and \ii{all}. Quantification is far from
the only place in natural language where complex functional meanings
are found, but it is a natural starting point, since it can be tested
in sentences whose structures are otherwise quite simple, and since it
has formed a standard case study in prior formal work on natural
language inference \cite{Icard:Moss:2013:LILT}.

@@

% TODO: Rewrite experimental setup.
We evaluate the models by randomly sampling 20\% of the data to
test on and training on the remainder. In this setting, the models are being asked to learn a
complete reasoning system for the limited language and logic presented
in the training data, but it is not being asked to generalize to test
examples that are substantially different from those it was trained
on. Crucially, though, to succeed on this task, the models must be able
to recognize all of the lexical relations between the nouns, verbs,
and quantifiers and how they interact. For instance, it might see
\eqref{p1} and \eqref{p2} in training and, from that information,
determine \eqref{p3}.

@@

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{??d RNN}  & \multicolumn{2}{c}{??d RNTN}\\
    \midrule
    Test & 35.4 &(7.5) &	99.1&(99.2)& \textbf{99.4...} & \textbf{(99.2)}\\
    Train & 35.4 &(7.5) &	-&	-&	-&	-\\	
    \bottomrule
  \end{tabular}
  \caption{Performance on the quantifier experiments.}
  \label{qresultstable}
\end{table} 
%
% do not allow a blank line --- adds too much space
%
\paragraph{Results} The results, reported in Figure \ref{qresultstable}, show that both models are able to learn to generalize the underlying logic well.

% TODO: More to say?

>>>>

We have seen that recursive NNs can learn an approximation of propositional
logic.  However, natural languages can express functional meanings of
considerably greater complexity than this.  As a step towards
investigating whether our models can capture this complexity, we now
directly measure the degree to which they are able to
develop suitable representations for the semantics of natural language
quantifiers like \ii{most} and \ii{all}. Quantification is far from
the only place in natural language where complex functional meanings
are found, but it is a natural starting point, since it can be tested
in sentences whose structures are otherwise quite simple, and since it
has formed a standard case study in prior formal work on natural
language inference \cite{Icard:Moss:2013:LILT}.

@@

We evaluate the models by randomly sampling 20\% of the data to
test on and training on the remainder. In this setting, each model being asked to learn a
complete reasoning system for the limited language and logic presented
in the training data. To succeed on this task, the model must be able
to recognize all of the lexical relations between the nouns, verbs,
and quantifiers and how they interact. For instance, it might see
\eqref{p1} and \eqref{p2} in training and, from that information,
determine \eqref{p3}.

@@

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{25d RNN}  & \multicolumn{2}{c}{25d RNTN}\\
    \midrule
    Test & 35.4 &(7.5) &	99.1&(99.2)& \textbf{99.5} & \textbf{(99.3)}\\
    Train & 35.4 &(7.5) &	100&		(100)&	100&	(100)\\	
    \bottomrule
  \end{tabular}
  \caption{\mynote{Performance on the quantifier experiments, given as \% correct and macroaveraged F1.}}
  \label{qresultstable}
\end{table} 
%
% do not allow a blank line --- adds too much space
%
\paragraph{Results} The results, reported in Figure \ref{qresultstable}, show that both models are able to learn to generalize the underlying logic well.

### 2014-12-02 writing/F'14 paper/recursion.tex
Paper update.

Complex embedding is a prominent feature of natural languages. 
The principle of compositionality explains how the meanings of 
such structures are derived systematically from their parts, thereby
 enabling people to use novel phrases and sentences.
Theoretical accounts of natural language syntax and semantics
therefore rely heavily on recursive structures, and we expect that
accurate computational models will have to come to grips with them as
well. The present section pursues this question for our two classes of
RNN. In our evaluations, we exploit the fact that our logical language
is infinite by testing on strings that are longer and more complex
than any seen in training.

@@

\begin{table}[tp]
  \centering\small
  \begin{subtable}[t]{0.45\textwidth}
    \centering
    \begin{tabular}[t]{l l}
      \toprule
      Formula     & Interpretation \\
      \midrule
      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
      \bottomrule
    \end{tabular}    
    \caption{Well-formed formulae. $\varphi$ and $\psi$
      range over all well-formed formulae, and $\sem{\cdot}$ is
      the interpretation function mapping formulae into $\{\True,
      \False\}$.}\label{tab:pl}
  \end{subtable}
  \begin{subtable}[t]{0.45\textwidth}
    \centering\vspace{4mm} % TODO: Is there a better command to ensure padding above the table?
    \begin{tabular}[t]{r c l}
      \toprule
      $\plneg a$        & $\natneg$ & $a$ \\
      $\plneg \plneg a$ & $\nateq$  & $a$ \\
      $a$               & $\natfor$ & $(a \plor b)$ \\
      $a$               & $\natrev$ & $(a \pland b)$ \\
      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
      $\plneg(\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
      \bottomrule
    \end{tabular}
    \caption{Examples of statements about relations between
      well-formed formulae, defined in terms of sets of satisfying
      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
  \end{subtable}
  \caption{Natural logic relations over sentences of propositional logic.}  
  % TODO: Regenerate figure from final model.
  \label{prop-figure}
\end{table}

% NOTE: There was some confusion at CSLI and the 224U lecture about
% whether this was theorem proving or some kind of model checking.
% Hopefully this is explicit enough.
Socher et al.~\shortcite{socher2012semantic} show that a matrix-vector RNN
model somewhat similar to our RNTN can learn a boolean logic, but the logic
discussed here is a much better approximation of the kind of
expressivity needed for natural language. In a boolean logic
 the atomic symbols are simply the
values $0$ and $1$, rather than variables over those values. While
learning the operators of that logic is not trivial, the outputs of
each operator can be represented accurately by a single bit. The
statements of propositional logic learned here describe conditions on
unknown truth values of propositions. As opposed to the two-way
contrasts seen in \newcite{socher2012semantic}, this logic distinguishes
between $2^{6} = 64$ possible assignments of truth values, and
expressions of this logic define arbitrary conditions on these
possible assignments, for a total of $2^{64}$ %($\approx 10^{20}$)
possible statements that the intermediate vector representations need
to be able to distinguish if we are to learn this logical system in its
full generality.

@@

\paragraph{Results} 
We trained both the RNN and RNTN models on data with up to 4
connectives (65k pairs) and tested it on examples of up to 12
connectives (44k pairs). We initialized the model parameters randomly,
including the vector representations of the six variables. The results
are shown in Figure~\ref{prop-results}. \mynote{In tuning, we found that the
RNN model was approximately optimal with 45-dimensional vector
representations, and the RNTN model was approximately optimal with 25
dimensions. We fixed the size of the feature vector for the classifier
at 75 dimensions. We found that the RNTN model was able to perform
almost perfectly on unseen small test examples, with accuracy above
98.7\% on formulae below length four (macro-F1 was above 95.5).
Starting at size four, performance gradually falls with increasing
size.  The RNN model did not perform well, reaching only 90.5\% (81.4
micro-F1) accuracy on the smallest test examples, and declining from
there to near-baseline performance on formulae of length 12. Training
accuracy was 99.6\% for the RNTN and 87.6\% for the RNN.}

The performance of the RNTN model on small unseen test examples
indicates that it learned a correct approximation of the underlying
logic. It appears that this approximation is accurate enough to yield
correct answers when the composition layer is only applied a small
number of times, but that the error in the approximation grows with
increasing expression size, resulting in gradually dropping performance. In
contrast, the RNN model did not appear to be able to accurately handle statements of any length, but it
appears that the approximation that the RNN did learn decayed similarly
with expression size.  This is not necessarily a significant flaw in the model, since it remains possible that  different training regimes would allow it to learn accurate representations of the larger expressions.

In an effort to better understand why the RNTN outperformed the plain
RNN, we evaluated both models on pairs of long formulae involving
binary connectives, to assess how well they distinguish representations for semantically distinct
formulae. We found that only the RNTN is able to learn
substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
 and $(a \pland (a \pland b))$ when the difference between the two is placed under multiple operators.
The RNN separates the bare symbols by a euclidean distance of 3.5, but this falls to less than one once the two are placed at under three operators, and quickly approaches zero as depth increases.
 The RNTN separates the bare symbols by 2.5, and this falls off much more gradually to 1.3 with twelve operators.

>>>>

Complex embedding is a prominent feature of natural languages. 
The principle of compositionality explains how the meanings of 
such structures are derived systematically from their parts, thereby
 enabling people to use novel phrases and sentences.
Theoretical accounts of natural language syntax and semantics
therefore rely heavily on recursive structures. This section examines the 
question of whether our models are able to learn a 
compositional semantics over such recursive structures.
In our evaluations, we exploit the fact that our logical language
is infinite by testing on strings that are longer and more complex
than any seen in training.

@@

\begin{table}[tp]
  \centering\small
  \begin{subtable}[t]{0.45\textwidth}
    \centering
    \begin{tabular}[t]{l l}
      \toprule
      Formula     & Interpretation \\
      \midrule
      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
      \bottomrule
    \end{tabular}    
    \caption{Well-formed formulae. $\varphi$ and $\psi$
      range over all well-formed formulae, and $\sem{\cdot}$ is
      the interpretation function mapping formulae into $\{\True,
      \False\}$.}\label{tab:pl}
  \end{subtable}
  \begin{subtable}[t]{0.45\textwidth}
    \centering\vspace{4mm}
    \begin{tabular}[t]{r c l}
      \toprule
      $\plneg a$        & $\natneg$ & $a$ \\
      $\plneg \plneg a$ & $\nateq$  & $a$ \\
      $a$               & $\natfor$ & $(a \plor b)$ \\
      $a$               & $\natrev$ & $(a \pland b)$ \\
      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
      $\plneg(\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
      \bottomrule
    \end{tabular}
    \caption{Examples of statements about relations between
      well-formed formulae, defined in terms of sets of satisfying
      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
  \end{subtable}
  \caption{Natural logic relations over sentences of propositional logic.}  
  \label{prop-figure}
\end{table}

Socher et al.~\shortcite{socher2012semantic} show that a matrix-vector RNN
model somewhat similar to our RNTN can learn boolean logic, 
a logic where the atomic symbols are simply the
values $0$ and $1$. While learning the operators of that logic is not trivial, the outputs of
each operator can be represented accurately by a single bit.
In the propositional logic used here, the atomic symbols are variables over those values. 
Statements of this logic thus describe conditions on
unknown truth values of propositions. This logic distinguishes
between $2^{6} = 64$ possible assignments of truth values, and
expressions of this logic define arbitrary conditions on these
possible assignments, for a total of $2^{64}$ %($\approx 10^{20}$)
possible statements that the intermediate vector representations need
to be able to distinguish if we are to learn this logical system in its
full generality.

@@

\paragraph{Results} 
We trained both the RNN and RNTN models on data with up to 4
connectives (65k pairs) and tested it on examples of up to 12
connectives (44k pairs). We initialized the model parameters randomly,
including the vector representations of the six variables. The results
are shown in Figure~\ref{prop-results}. \mynote{In tuning, we found that the
RNN model was approximately optimal with 45-dimensional vector
representations, and the RNTN model was approximately optimal with 25
dimensions. We fixed the size of the feature vector for the classifier
at 75 dimensions. We found that the RNTN model was able to perform
almost perfectly on unseen small test examples, with accuracy above
98.7\% on formulae below length four (macro-F1 was above 95.5).
Starting at size four, performance gradually falls with increasing
size.  The RNN model did not perform well, reaching only 90.5\% (81.4
macro-F1) accuracy on the smallest test examples, and declining from
there to near-baseline performance on formulae of length 12. Training
accuracy was 99.6\% for the RNTN and 87.6\% for the RNN.}

The performance of the RNTN model on small unseen test examples
indicates that it learned a correct approximation of the underlying
logic. It appears that this approximation is accurate enough to yield
correct answers when the composition layer is only applied a small
number of times, but that the error in the approximation grows with
increasing expression size, resulting in gradually dropping performance. In
contrast, the RNN model did not appear to be able to accurately handle statements of any length, but it
appears that the approximation that the RNN did learn decayed similarly
with expression size.

\mynote{In an effort to better understand why the RNTN outperformed the plain
RNN, we evaluated both models on pairs of long formulae involving
binary connectives, to assess how well they distinguish representations for semantically distinct
formulae. We found that only the RNTN is able to learn
substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
 and $(a \pland (a \pland b))$ when the difference between the two is placed under multiple operators.
The RNN separates the bare symbols by a euclidean distance of 3.5, but this falls to less than one once the two are placed under three operators, and quickly approaches zero as depth increases.
 The RNTN separates the bare symbols by 2.5, and this falls off much more gradually to 1.3 with twelve operators.}

 % TODO: If results are still bad tomorrow, write something about why this is not fatal to the program.

### 2014-12-02 writing/F'14 paper/sick.tex
Minor paper updates.
Paper update.

The tree-pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference in the face of the noisy labels and diverse range of linguistic structures seen in typical natural language data. To ensure that this is the case, we also train a version of our model on an existing competition dataset, the SICK textual entailment challenge \cite{marelli2014sick}. The corpus consists of about 10k natural language sentence pairs (generated using data expansion from a smaller set), labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}. At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest hand labeled entailment corpus that we are aware of.

@@

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ } r@{ \ } r@{ \ } }
    \toprule
    ~&\multicolumn{1}{c}{$\natind$ only} & \multicolumn{1}{c}{??d RNN}  & \multicolumn{1}{c}{??d RNTN}\\
    \midrule
    Test & 56.7 &	...& \textbf{75.5} \\
    Train &- &- &-  \\
    \bottomrule
  \end{tabular}
  \caption{\mynote{Performance on SICK.}}
  \label{sresultstable}
\end{table}

Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships (especially exclusion, as in \ii{cat}--\ii{dog}), it is possible to train our model to perform competitively on textual entailment. Our performance did not reach the state of the art (84.6\%), but exceeded the median submission to the competition (77.06\%) in a field including many models which used sophisticated hand-engineered features specific to the specific version of the entailment task at hand.

Table \ref{examplesofsickdata} shows a few typical examples of correct classifications. The model appears to be able to successfully recognize negation and the insertion and deletion of words, as demonstrated in the first two examples. The latter two show that the model still recognize entailments and contradictions between sentences which are not superficially similar, such as passive--active pairs and sentences with low lexical overlap.

>>>>

The tree pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference in the face of the noisy labels and diverse range of linguistic structures seen in typical natural language data, and our experiments on artificial data are not alone sufficient to show this. To investigate the ability of the model to learn on natural data, we train a version of our model on the SICK textual entailment challenge corpus \cite{marelli2014sick}. The corpus consists of about 10k natural language sentence pairs (generated using data expansion from a smaller set), labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}. At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest hand labeled entailment corpus that we are aware of.

@@

% TODO: JMLR version of hinton paper.
% TODO: Add Karpathy to TACL paper

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ } r@{ \ } r@{ \ } }
    \toprule
    ~&\multicolumn{1}{c}{$\natind$ only} & \multicolumn{1}{c}{50d RNN}  & \multicolumn{1}{c}{50d RNTN}\\
    \midrule
    Test & 56.7 &	...& \textbf{75.5} \\
    Train &- &- &-  \\
    \bottomrule
  \end{tabular}
  \caption{\mynote{Performance on SICK. Training figures include performance on the Denotation Graph data.}}
  \label{sresultstable}
\end{table}

Despite the small amount of high quality training data available and the lack of resources for learning lexical relationships (especially exclusion, as in \ii{cat}--\ii{dog}), it is possible to train our model to perform competitively on textual entailment. Our performance did not reach the state of the art (84.6\%), but exceeded the median submission to the competition (77.06\%) in a field including many models which used sophisticated hand-engineered features and lexical resources specific to the version of the entailment task at hand.

Table \ref{examplesofsickdata} shows a few typical examples of correct classifications. The model is able to successfully recognize negation and the insertion and deletion of words, as demonstrated in the first two examples. The latter two show that the model can recognize entailments and contradictions between sentences which are not superficially similar, such as passive--active pairs and sentences with low lexical overlap.
